### 2021-02-04 examples/Toy-example-balanced-accessibility.Rmd

---
title: "R Notebook"
output: html_notebook
---

Load libraries:
```{r message=FALSE}
library(data.table)
library(dplyr)
library(gdistance)
library(kableExtra)
library(raster)
library(rgdal)
library(sf)
library(tidyverse)
library(units)
```

# Load data

Load geography:
```{r}
load("population_sobi.RData") #interpolated population in SoBi area
load("sobi_hubs.RData") #location of SoBi hubs, minus Van Wagner's beach
load("sobi_service.RData") #core service area for SoBi
load("ttm_walk.RData") #travel time matrix for interpolated population in SoBi area
```

```{r}
summary(ttm_walk) #time is in minutes
```

# Check the data

Plot the interpolated population:
```{r}
ggplot(data = population_sobi) +
  geom_sf(aes(color = population)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1)
```

Notice that the population units begin with UID 1 and ends with UID 11388:
```{r}
summary(population_sobi)
```

The travel time table, on the other hand, begins with population UID 15 and ends with UID 11305:
```{r}
summary(ttm_walk)
```

What explains this discrepancy? We can compare the population units that are in the population dataset but not in the travel time table:
```{r}
ggplot() +
  geom_sf(data = sobi_service) +
  geom_sf(data = population_sobi %>%
            filter(!(UID %in% (distinct(ttm_walk, # filter observations with UIDs that are present in the travel time table
                                        UID) %>% 
                                 pull(UID)))))

```

The reason that some population units do not appear in the travel time table is that they are outside of the service area of SoBi and/or cannot reach SoBi hubs by walking.

# Threshold

Calculate quantiles according to distance. This is done to select a threshold:
```{r}
quantiles_d <- quantile(as.vector(ttm_walk$travel_time), probs = c(0, 1/10, 3/10, 5/10, 7/10, 8/10, 9/10))

dist_threshold <- quantile(as.vector(ttm_walk$travel_time), probs = c(0, 1/10, 3/10, 5/10, 7/10, 8/10, 9/10))[[5]]
```

The distance threshold is equivalent to the _kernel bandwidth_ or _catchment distance_.

# Set up matrix

Add population information:
```{r}
ttm_walk <- left_join(ttm_walk, 
                      population_sobi,
                      by = "UID")
```

Drop the geometry:
```{r}
ttm_walk <- ttm_walk %>% dplyr::select(-geometry)
```

Add number of bicycle racks per SoBi hub as a measure of the level of supply:
```{r}
ttm_walk <- left_join(ttm_walk, sobi_hubs %>% dplyr::select(OBJECTID, RACKS_AMOU), by=c("OBJECTID" = "OBJECTID"))
ttm_walk <- ttm_walk %>% dplyr::select(-geometry)
head(ttm_walk)
```

Rename columns:
```{r}
names(ttm_walk) <- c('UID', 'hub', 'travel_time', 'population', 'racks')
```

Calculate impedance (binary):
```{r eval=TRUE}
# Run only if binary function desired
# Change chunk option eval=TRUE to FALSE if impedance is Gaussian function
ttm_walk <- ttm_walk %>%
  mutate(impedance_binary = ifelse(travel_time <= dist_threshold, 1, 0))
```

Check the value of the dist_threshold:
```{r}
dist_threshold
```

Our calculations assume that people do not go further than 24 min to reach a SoBi hub.

Calculate impedance (Gaussian):
```{r eval=FALSE}
# Run only if Gaussian function desired
# Change chunk option eval=FALSE to TRUE if impedance is Gaussian function
ttm_walk <- ttm_walk %>%
  mutate(impedance_gaussian = case_when(travel_time <= quantiles_d[1] ~ 0.945,
                                        (travel_time > quantiles_d[1]) & (travel_time <= quantiles_d[3]) ~ 0.600,
                                        (travel_time > quantiles_d[3]) & (travel_time <= quantiles_d[5]) ~ 0.242,
                                        travel_time > quantiles_d[5] ~ 0))
```

Calculate the sum of the impedance by population for the balancing factors. In the binary case, this should be the same as the number of SoBi hubs that each interpolated population unit can reach; when the impedance is not binary, this is the sum of all impedance values:
```{r}
sum_b1 <- ttm_walk %>%
  group_by(UID) %>%
  summarize(sum_b1 = sum(impedance_binary),
            .groups = "drop")
sum_b1
```

The interpolated population at unit 15 can reach eight SoBi hubs, the population at unit 16 can reach six SoBi hubs, the population at unit 30 can reach zero hubs clinics, etc..

Sanity check: how many stations are within 24 minutes or less from population UID 15?
```{r}
ttm_walk %>%
  filter(UID == 15, travel_time <= 24) %>%
  nrow()
```

Calculate the sum of the impedance by SoBi hubs for the balancing factors. In the binary case, this should be the same as the number of interpolated population unit that each SoBi hub serves; when the impedance is not binary, this is the sum of all impedance values:
```{r}
sum_b2 <- ttm_walk %>%
  group_by(hub) %>%
  summarize(sum_b2 = sum(impedance_binary))
sum_b2
```

Hub 1 serves 1502 interpolated population units, hub 2 serves 1393 population units, etc.

Sanity check: how many population units are served by hub 1?
```{r}
ttm_walk %>%
  filter(hub == 1, travel_time <= 24) %>%
  nrow()
```

The values check.

Next, join the sum of the impedance to the table to calculate the balanced impedance values:
```{r}
ttm_walk <- ttm_walk %>%
  left_join(sum_b1, by = "UID") %>%
  left_join(sum_b2, by = "hub")
```

Check `sum_b1`:
```{r}
ttm_walk$sum_b1 %>% 
  summary()
```

**Important:** notice that some values of `sum_b1` are zeros! This means that some population units cannot reach a SoBi hub in at most 24 min. Since the balancing impedance is calculated by diving the impedance by `sum_b1`, this will lead to NaNs (divisions by zero). To avoid issues we will remove any population units that reach zero hubs:
```{r}
ttm_walk <- ttm_walk %>%
  filter(sum_b1 > 0)
```

Are there any hubs that serve zero population units? (same issue with division by zero):
```{r}
ttm_walk$sum_b2 %>% 
  summary()
```

No zeros or other strange values, we're good.

Calculate balanced impedance values:
```{r}
ttm_walk <- ttm_walk %>%
  mutate(balanced_impedance_1 = impedance_binary/sum_b1,
         balanced_impedance_2 = impedance_binary/sum_b2)
```

# Calculate level of service

Check the balanced_impedance factors for the population units:
```{r}
ttm_walk %>% 
  dplyr::filter(hub == "46") %>%
  pull(balanced_impedance_2) %>%
  sum()
```

The sum should be 1.

[**Note: not sure if this step is needed for SoBi?** **Note for Elise: this is only an intermediate check in the analysis, it is not strictly necessary**] *Notice that the population of each center is sorted among the SoBi hubs proportionally. Thanks to this property, the sum of the population serviced equals the total population in the system:*
```{r}
ttm_walk %>%
  group_by(UID) %>%
  summarize(pop_served = sum(sum(population * balanced_impedance_1)),
            .groups = "drop") %>%
  pull(pop_served) %>%
  sum()

population_sobi %>% 
  filter(UID %in% (distinct(ttm_walk,
                             UID) %>%
             pull(UID))) %>%
  pull(population) %>% 
  sum()
```

The level of service of each hub is the number of bicycle racks at the hub, divided by the population that they serve:
```{r}
sobi_los <- ttm_walk %>%
  dplyr::filter(impedance_binary > 0) %>% # filter impedance == 0 to avoid divisions by zero
  group_by(hub) %>%
  summarize(sobi_los = first(racks) / sum((population * balanced_impedance_1)),
            .groups = "drop")
sobi_los
```

Hub 1 has 0.01538 bicycles per person, hub 2 has 0.02446 bicycles per person, etc.

Join the level of service to the table:
```{r}
ttm_walk <- ttm_walk %>%
  left_join(sobi_los, by = "hub")
```

# Calculate accessibility

To calculate accessibility, the level of service of each hub needs to be distributed proportionally to the population units. 
```{r}
sobi_accessibility <- ttm_walk %>%
  group_by(UID) %>%
  summarize(accessibility = sum(sobi_los * balanced_impedance_2),
            .groups = "drop")
sobi_accessibility
```

Check that the sum of the accessibility is equal to the available level of service:
```{r}
sobi_accessibility %>% 
  pull(accessibility) %>% 
  sum()

sobi_los %>% 
  pull(sobi_los) %>%
  sum()
```

```{r}
ggplot(data = population_sobi) +
  geom_sf(aes(color = population))
```

Join the accessibility to the population units:
```{r}
population_sobi <- population_sobi %>%
  left_join(sobi_accessibility,
            by = "UID")
```

Plot accessibility:
```{r}
ggplot(data = population_sobi) +
  geom_sf(aes(color = accessibility)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_distiller(palette = "YlOrRd", 
                        direction = 1)
```

[**END FOR SOBI ACCESSIBILITY**]

<!--The text that follows was mostly copied from PLoS ONE paper-->

## Disparity Analysis

An advantage of the use of adjusted weights for proportional allocation of demand and level of service it that, after rectifying the inflation artifact, they make it is possible to conduct accessibility disparity analysis in a very intuitive way.

For instance, an analyst interested in equity analysis could allocate the total level of service uniformly to every population unit. In other words, the total level of service (which equals the sum of accessibility over the system) can be divided by the number of population units in the system to return the Average Local Population Unit PPR. The resulting mean value, call it $L_i^e$ then would be assigned to the population units as their "equitable" share of the total level of service in the system. Next, the equitative distribution of the level of service in each population unit is substracted from the estimated accessibility to arrive at a disparity index. When the difference between these two quantities is positive, this would indicate that accessibility at that location exceeds its equitable share of level of service. On the other hand, when the difference is negative, accessibility at the location is below its equitable share of the level of service.

This approach is reminiscent of the Spatial Access Ratio (SPAR) proposed by Wan et al. [-@Wan2012SPAR], which is calculated as the ratio between a population center's accessibility and the mean accessibility across all population centers. Wan et al. [-@Wan2012] calculate SPAR based on the results of their 3SFCA method, by rescaling the accessibility measures to reflect the percentage difference in each population center's accessibility relative to the mean. This measure is designed to overcome the sensitivity of existing FCA metrics to the impedance function. In contrast, the approach proposed here, enables more intuitive and interpretable results by preserving the system-wide population and level of service. In this way, a disparity index is useful to highlight the absolute difference in accessible provider-to-population ratios across population units.

The disparity for SoBi is shown in the following map. The figure reveals the spatial distribution in disparity, with levels of access that are lower than the mean in the periphery of the service area, compared to levels of access that are greater than the mean in the higher-density and more connected urban center.

```{r}
population_sobi %>%
  mutate(Lei = accessibility - sum(sobi_los$sobi_los)/nrow(population_sobi %>% drop_na(accessibility))) %>%
  ggplot() +
  geom_sf(aes(color = Lei)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_color_gradient2()
```

```{r accessibility_disparities, echo=FALSE, include=FALSE}
#Disparities in this case is defined as the difference between the calculated accessibility (which is interpreted as the local provider-to-population ratio) and the local provider-to-population ratio if the level of service was distributed uniformly
accessibility_disparities <- transmute(accessibility, DA, 
                                      Binary = Binary_Adjusted - sum(los_binary_adjusted)/nrow(accessibility_binary_adjusted),
                                      Stepwise = Stepwise_Adjusted- sum(los_stepwise_adjusted)/nrow(accessibility_stepwise_adjusted))
```

```{r join-geography, echo=FALSE, include=FALSE}
#Join accessibility measures to geography.
hamilton_da.sf_accessibility <- left_join(hamilton_da.sf, accessibility, by = "DA")
hamilton_da.sf_comparison <- left_join(hamilton_da.sf, accessibility_comparison, by = "DA")
hamilton_da.sf_disparities <- left_join(hamilton_da.sf, accessibility_disparities, by = "DA")
```

<!--
# Complete  (or tl;dr version)

The full algorithm can be written in this way.

Obtain the matrix in long form (this is line 85 in this file):
```{r}
long_matrix <- left_join(distance_matrix_long, Clinics_data, by=c("to" = "id"))
```

Calculate the accessibility and level of service:
```{r}
# Calculate impedance (binary)
# long_matrix <- long_matrix %>%
#   mutate(distance = drop_units(distance),
#          impedance = ifelse(distance <= dist_threshold, 1, 0))

# Calculate impedance (Gaussian)
long_matrix <- long_matrix %>%
  mutate(distance = drop_units(distance),
         impedance = case_when(distance <= quantiles_d[1] ~ 0.945,
                               (distance > quantiles_d[1]) & (distance <= quantiles_d[3]) ~ 0.600,
                               (distance > quantiles_d[3]) & (distance <= quantiles_d[5]) ~ 0.242,
                               distance > quantiles_d[5] ~ 0))

# Calculate first set of balancing factors
sum_b1 <- long_matrix %>%
  group_by(id) %>%
  summarize(sum_b1 = sum(impedance),
            .groups = "drop")

# Calculate second set of balancing factors
sum_b2 <- long_matrix %>%
  group_by(to) %>%
  summarize(sum_b2 = sum(impedance),
            .groups = "drop")

# Join the sum of the impedance to the table to calculate the balanced impedance values:
long_matrix <- long_matrix %>%
  left_join(sum_b1, by = "id") %>%
  left_join(sum_b2, by = "to")

# Calculate the balanced impedance
long_matrix <- long_matrix %>%
  mutate(balanced_impedance_1 = impedance/sum_b1,
         balanced_impedance_2 = impedance/sum_b2)

# Calculate the level of service
los <- long_matrix %>%
  filter(impedance > 0) %>% # filter impedance == 0 to avoid divisions by zero
  group_by(to) %>%
  summarize(los = first(Doctors) / sum((Population * balanced_impedance_1)),
            .groups = "drop")

# Join the level of service to the table:
long_matrix <- long_matrix %>%
  left_join(los, by = "to")

# Calculate accessibility
accessibility <- long_matrix %>%
  group_by(id) %>%
  summarize(accessibility = sum(los * balanced_impedance_2),
            .groups = "drop")
```

This is the accessibility:
```{r}
accessibility
```

This is the level of service:
```{r}
los
```

-->