### 2021-03-28 Bike-share-spatial-equity/Highlights.Rmd

---
title: "Highlights"
output:
  pdf_document: default
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

* Balanced floating catchment area methods are applied to measure accessibility to PBSP

* Contribution of equity stations to accessibility assessed at different thresholds for walking time to stations

* Equity stations increased both horizontal and vertical accessibility, albeit only modestly

* Lowest income group has lowest levels of accessibility, even with equity stations

* Locations for new stations are identified to increase equitable access to PBSP