### 2021-08-23 Bike-share-spatial-equity/Highlights.Rmd
Revised Highlights - June 2021

---
title: "Highlights"
output:
  pdf_document: default
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

* Balanced floating catchment area method applied to measure accessibility to bikeshare

* Contribution of stations added to increase equity assessed at different thresholds

* Added stations increased accessibility and reduced inequities, albeit only modestly

* Lowest income group has lowest levels of accessibility, even with "equity" stations

* Location and size of docking stations matters for increasing equity in accessibility