### 2021-10-14 Bike-share-spatial-equity/Untitled/Untitled.Rmd
Revised Manuscript - Round 2

---
title: "Exercise 1: Energy consumption trends"
author:
- name: My Name
  student_number: 123456789
subject: "ENVSOCTY 3LT3"
highlights: |
    Write a brief summary from your perspective of the highlights of working on this assignment.

    You can use more than one paragraph but remember to indent the paragraphs. Try to keep this at less than 200 (?) words.
threshold_concepts: 
# Write the concepts that in your opinion are threshold concepts in this exercise.
- threshold concept 1
- threshold concept 2 
- threshold concept 3
- threshold concept 4
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    # The project-template-default.tex file was heavily  adapted from Steven V. Miller's template for academic manuscripts. See:
    # http://svmiller.com/blog/2016/02/svm-r-markdown-manuscript/
    # https://github.com/svmiller/svm-r-markdown-templates/blob/master/svm-latex-ms.tex
    template: exercise-template-default.tex
always_allow_html: true
---

# Introduction

Transport is fraught with many issues including growth in demand for energy use, greenhouse gases, urban air pollution and noise, negative health impacts and road traffic accidents and fatalities. It is well argued that these problems are not typical to one region but are becoming global in nature. Similar trends have been observed in many places around the world, albeit there are still individual differences among various countries. The observed data suggest an emerging transport crisis that could potentially compromise our quality of life.

In this exercise, you will analyze trends pertaining to energy use and emissions in several countries. In so doing, you will gain a better understanding of emerging transportation trends and their impacts on quality of life.

* Readings:
    + Installation of R.pdf
    + 00 Overview of Data Analysis and Visualization Using R.Rmd (found under Labs/Assignments)
    + 01 Regression Analysis.Rmd (found under READINGS)
* Due date:	See course outline for due date.
* Materials: This R Notebook, and the following Excel file (found under Exercises)
    + `Energy and Emissions.xlsx`

# Instructions

Answers must be provided by typing your solutions using RStudio and this R Notebook file. Change the name of the file to `Exercise 1 (Name-Student Number).Rmd`

The variables in the data file are:

Variable | Definition
-- | --
Country	| Country name
Area | Surface land area of country in sq.km
Population | Population
PYear | Year of estimated population
GDPPC | GDP per capita (nominal) in US dollars
bblpd | barrels of oil per day
EYear | Year of estimated usage of barrels of oil per day
CO2_1995 | CO2 emissions in kilotonnes per year (1995)
CO2_2005 | CO2 emissions in kilotonnes per year (2005)
CO2_2015 | CO2 emissions in kilotonnes per year (2015)

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*. When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

# Preliminaries

Again, you want to clear the working space to make sure that you do not have extraneous items there when you begin your work. The command in R to clear the workspace is `rm` (for "remove"), followed by a list of items to be removed. To clear the workspace from _all_ objects, do the following:

```{r}
rm(list = ls())
```

Begin by loading the datasets that you will need for this exercise. They are available from the Readings section. Note that you also need the package `readxl`. Make sure that it is installed before loading it with `library()`.

```{r}
library(envsocty3LT3) # Course package
library(ggplot2)
```

# Context

Words go here.

# Questions

## Energy Use and Emission Trends (100 marks)

1.	What was the approximate population of Earth circa 2016? How many barrels of oil per day did this population consume? (10)

2.	What proportion of the world's oil is consumed by Canada and the US? Do these two countries have the same per capita levels of demand for energy use? Why do you think this is so? (15)

3.	Calculate the consumption of oil in barrels per day per 1000 people for the countries listed in the data frame. What type of pattern do you observe? Try to explain the pattern. (20)

Hint: rank countries according to their oil consumption per 1000 people. You can use the `order()` function to do this. A good explanation is at the following link:

http://www.statmethods.net/management/sorting.html.

4. Use simple regression analysis to assess the relationship in the dataset between a country's per-capita CO2 emissions (for 2015) and per-capita oil consumption. Start by creating a scatterplot of these two variables; then add labels to your plot (try changing the `size` of the text for clarity).

What does a linear regression say is the relationship between per-capita CO2 emissions and per-capita oil consumption? How significant is the relationship, according to a regression? Which countries are cleaner and which countries are less clean in their use of energy? (25)

5. How much CO2 would be produced, if every person in the world produced CO2 equal to that produced by a Canadian? What if every person in the world produced CO2 equal to that produced by a Chinese citizen in 2015? Compare both results to the actual CO2 production of the world in 2015 and use a linear regression to explain the relationship. (15)

6.	Using the data in the data frame, find the six largest CO2 emitters in 1995, and then calculate the increase or decrease in the amount of generated CO2 emissions for each of them in the time period 1995 to 2015. Compare the size of the numbers. Which of these countries saw the largest increase in emissions since 1995? Why do you think so? (10)

7.	Which country in the world had the highest CO2 emissions per square kilometer of land area in 2015? Which country has the lowest? (5)